---
title: "hw3_grant"
author: "Grant King"
date: "2/18/2020"
output: html_document
---

```{r}
rm(list=ls())
library(car)
library(caret)
library(tidyr)
library(dplyr)
library(stats)
library(readxl)
library(ggplot2)
library(clipr)
library(ks)
library(EnvStats)
library(triangle)
library(knitr)
library(MASS)
library(rgl)
library(graphics)
library(truncnorm)
library(scales)
library(extraDistr)
```


Setting up distributions and calculating the probability of a producing well
```{r}

num_simulations = 1000000
set.seed(1)

hydrocarbons <- rtruncnorm(n = num_simulations, a = 0, b = 1, mean = .99, sd = .05)
structure <- rnorm(n=num_simulations, mean=1, sd=0)
reservoir <- rtruncnorm(n = num_simulations, a = 0, b = 1, mean = .80, sd = .10)
seal <- rnorm(n=num_simulations, mean=1, sd=0)

total_prob <- hydrocarbons * structure * reservoir * seal
```

Plotting these probability distributions

```{r}
# plot hydrocarbons and reservoir histograms
  ggplot(as.data.frame(hydrocarbons), aes(x=hydrocarbons)) +
    geom_histogram(col = 'black',
                  fill = 'orange',
                  bins = 40) +
    geom_vline(aes(xintercept=mean(hydrocarbons)),
                  color="darkred",
                  linetype="dashed",
                  size=1) +
    labs(x="Probability of Hydrocarbons Present", y="Frequency") +
    theme(panel.background = element_rect(color = 'lightgrey'),
                  panel.grid.major.y = element_line(color = 'white'),
                  panel.grid = element_blank()
  )
  
  ggplot(as.data.frame(reservoir), aes(x=reservoir)) +
    geom_histogram(col = 'black',
                  fill = 'orange',
                  bins = 40) +
    geom_vline(aes(xintercept=mean(reservoir)),
                  color="darkred",
                  linetype="dashed",
                  size=1) +
    labs(x="Probability of Developed Reservoir", y="Frequency") +
    theme(panel.background = element_rect(color = 'lightgrey'),
                  panel.grid.major.y = element_line(color = 'white'),
                  panel.grid = element_blank()
  )
    
  '---------------------------------------------------------------------------------------------------'
  
  # plot histogram of wet well probability
  ggplot(as.data.frame(total_prob), aes(x=total_prob)) +
    geom_histogram(col = 'black',
                  fill = 'orange',
                  bins = 40) +
    geom_vline(aes(xintercept=mean(total_prob)),
                  color="darkred",
                  linetype="dashed",
                  size=1) +
    labs(x="Probability of Producing Well", y="Frequency") +
    theme(panel.background = element_rect(color = 'lightgrey'),
                  panel.grid.major.y = element_line(color = 'white'),
                  panel.grid = element_blank()
  )
```


Calculating distribution for the number of wells and number of producing wells in the project

```{r}

  num_wells <- rdunif(n = num_simulations, min = 10, max = 30)
  producing_wells <- rbinom(n=num_simulations, size = num_wells, prob = total_prob)
  proportion_wet_wells <- producing_wells / num_wells
  
  
  
  ggplot(as.data.frame(proportion_wet_wells), aes(x=proportion_wet_wells)) +
    geom_histogram(col = 'black',
                  fill = 'orange',
                  bins = 30) +
    geom_vline(aes(xintercept=mean(proportion_wet_wells)),
                  color="darkred",
                  linetype="dashed",
                  size=1) +
    labs(x="Porportion of Wet Wells", y="Frequency") +
    theme(panel.background = element_rect(color = 'lightgrey'),
                  panel.grid.major.y = element_line(color = 'white'),
                  panel.grid = element_blank()
  )
  
  
  
  
  dry_wells <- num_wells - producing_wells

  hist(producing_wells, breaks = 50)
  hist(dry_wells, breaks=50)
  

  
```


Bringing back in the calculations from HW2 to calculate the distribution for NPV. For each well in the project, we will need to calculate NPV if it is a wet well, or drilling cost if it is a wet well.

This will run all of the hw1 code to get the data/objects. 
Paste the path to 'hw1_script.R' below..  
```{r setup, echo=FALSE}
opts_chunk$set(echo = FALSE, cache=FALSE)
read_chunk('C:/Users/grant/Documents/MSA/Spring/sim_risk/hw_2/hw1_script.R')
```

click green arrow to run this chunk
```{r hw1_results}
```

Dry Well Cost
```{r}
drill.cost.dry <- nonnorm_cost_2020 * 1000
lease.cost.dry <- rnorm(n=num_simulations * 20, mean = 600, sd = 50) * 960
seismic.cost.dry <- rnorm(n=num_simulations * 20, mean = 3, sd = .35) * 43000
prof.cost.dry <- rtriangle(n=num_simulations * 20, a = 172000, b = 279500, c = 215000)

dry.well.cost <- drill.cost.dry + lease.cost.dry + seismic.cost.dry + prof.cost.dry

```

Wet Well NPV

```{r}

#initial cost
drill.cost <- nonnorm_cost_2020 * 1000
lease.cost <- rnorm(n=num_simulations * 20, mean = 600, sd = 50) * 960
seismic.cost <- rnorm(n=num_simulations * 20, mean = 3, sd = .35) * 43000
completion.cost <- rnorm(n=num_simulations * 20, mean = 390000, sd = 50000)
prof.cost <- rtriangle(n=num_simulations * 20, a = 172000, b = 279500, c = 215000)

initial.cost <- drill.cost + lease.cost + seismic.cost + completion.cost + prof.cost


#revenue and stuff
volume_cumulative_30yrs <- rep(0,num_simulations * 20)
R <- matrix(data=cbind(1,0.64, 0.64, 1), nrow=2)
U <- t(chol(R))

standardize <- function(x){
  x.std = (x - mean(x))/sd(x)
  return(x.std)
}

destandardize <- function(x.std, x){
  x.old = (x.std * sd(x)) + mean(x)
  return(x.old)
}

set.seed(1)

IP.r <- rlnorm(n=num_simulations * 20, mean= 6, sd = 0.28)
decline.r <- runif(n=num_simulations * 20,min=.15, max=.32)
Both.r <- cbind(standardize(IP.r), standardize(decline.r))
IPD.r <- U %*% t(Both.r)
IPD.r <- t(IPD.r)
  
final.IP.r <- destandardize(IPD.r[,1], IP.r)
final.decline.r <- destandardize(IPD.r[,2], decline.r)

#changing decline to be within its regular bounds of .15 to .32, even though this will slightly mess with correlation
final.decline.r[final.decline.r<.15] <- .15
final.decline.r[final.decline.r>.32] <- .32


rate0 <- final.IP.r

#simulating declining rates for 30 years
for(year in 1:15){
  #set rate at end of year equal to rate at beginning minus decline
  #if the decline is more than 100% (due to correlation stuff) set rate=0
  rate1 <- rate0*(1 - final.decline.r)
  #calculate volume of oil in year
  eval(parse(text=paste('volume_',year,' <- 365 * (rate0 + rate1) / 2',sep="")))
  #set rate at beginning of next year to rate at end of previous year
  rate0 <- rate1
}


for (year in 1:15){
  eval(parse(text=paste('price_',year,' <- rtriangle(n=num_simulations * 20, a=projections[[',year+1,',3]], b=projections[[',year+1,',2]], c=projections[[',year+1,',4]])',sep="")))
  }
  
NRI <- rnorm(n=num_simulations * 20, mean=.75, sd=.02)
for (year in 1:15){
  eval(parse(text=paste('rev_',year,' <- (volume_',year,' * price_',year,') * NRI * (1-.046)', sep="")))
}
  
for (year in 1:15){
  eval(parse(text=paste('op.cost_',year,' <- rnorm(n=num_simulations * 20, mean=2.25, sd=0.3)', sep="")))
  eval(parse(text=paste('op.cost.total_',year,' <- op.cost_',year,' * volume_',year, sep="")))
  eval(parse(text=paste('cost.total_',year,' <- op.cost.total_',year,' + prof.cost', sep="")))
}
  
for (year in 1:15){
  eval(parse(text=paste('fnr_',year,' <- rev_',year,' - cost.total_',year, sep="")))
}
  
NPV = -initial.cost +
  fnr_1/(1.1) +
  fnr_2/(1.1^2) +
  fnr_3/(1.1^3) +
  fnr_4/(1.1^4) +
  fnr_5/(1.1^5) +
  fnr_6/(1.1^6) +
  fnr_7/(1.1^7) +
  fnr_8/(1.1^8) +
  fnr_9/(1.1^9) +
  fnr_10/(1.1^10) +
  fnr_11/(1.1^11) +
  fnr_12/(1.1^12) +
  fnr_13/(1.1^13) +
  fnr_14/(1.1^14) +
  fnr_15/(1.1^15)
```


```{r}
  # simulate NPV for entire project (dry and wet well count TBD from above)
  total_NPV <- rep(0,num_simulations) 
  i_prod <- 1
  i_dry <- 1
  for (i in 1:num_simulations){
    #adding the NPV of a number of observations in the distribution equal to the number of producing wells
    total_NPV[i] <- total_NPV[i] + sum(NPV[i_prod:i_prod+producing_wells[i]])
    #subtracting the dry.well.cost of a number of observations in the distribution equal to the number of dry wells
    total_NPV[i] <- total_NPV[i] - sum(dry.well.cost[i_dry:i_dry+dry_wells[i]])
    #iterating the counters for the producing well and dry well distributions so samples aren't repeated
    i_prod <- i_prod + producing_wells[i]
    i_dry <- i_dry + dry_wells[i]
  }
  
  '---------------------------------------------------------------------------------------------------'

  # calculate expected return including 5% VaR and 5% CVaR and interpret these values

expected_return <- mean(total_NPV)
VaR <- quantile(total_NPV, .05)
cVaR <- mean(total_NPV[total_NPV < VaR])


print(paste("Expected Return:", expected_return))
print(paste("95% Value at Risk:", VaR))
print(paste("95% Conditioned Value at Risk", cVaR))


  # plot histogram of wet well probability
  ggplot(as.data.frame(total_NPV), aes(x=total_NPV)) +
    geom_histogram(col = 'black',
                  fill = 'orange',
                  bins = 40) +
    geom_vline(aes(xintercept=mean(total_NPV)),
                  color="darkred",
                  linetype="dashed",
                  size=1) +
    labs(x="Net Present Value of Project", y="Frequency") +
    theme(panel.background = element_rect(color = 'lightgrey'),
                  panel.grid.major.y = element_line(color = 'white'),
                  panel.grid = element_blank()) + 
    scale_x_continuous(labels = comma)
  
```