---
title: "Sim and Risk -- HW1"
author: "Will Elmore"
date: "1/30/2020"

---

```{r}
library(car)
library(caret)
library(tidyr)
library(dplyr)
library(stats)
library(readxl)
library(ggplot2)
library(clipr)
library(EnvStats)
library(triangle)
```




Importing - NOTE: I cleaned the data manually to have shorter column names and to remove the top two rows of each sheet (which is metadata)

column names for the costs df were changed to
["Date",
 "crude_well_cost",
 "gas_well_cost",
 "dry_well_cost",
 "crude_return",
 "gas_return",
 "dry_return"]
```{r}
# Importing projections
projections <- read_excel("Analysis_Data.xlsx", sheet = "Price Projections")

# importing drilling costs
costs <- read_excel("Analysis_Data.xlsx", sheet = "Drilling Cost")


```


1) Build a kernel density estimate of the distribution of changes between 1990 and 2006. Use this KDE to simulate average changes from 2006-2012 as well.


subsetting 1990:2006 
```{r}
costs.sub <- costs[32:47,]
costs.sub[,2:7] <- sapply(costs.sub[,2:7], as.numeric)
```

"Instead of focusing on costs for oil, gas, and dry wells individually, the Company’s
analysts recommend to treat them all equally and assume an average cost applies to them all."

Concatenating the different columns so that we have 48 total observations
  - bandwith is 0.07935
```{r}
costs_vec <- c(costs.sub[['crude_return']], costs.sub[['gas_return']], costs.sub[['dry_return']])
costs.sub$avg_costs <- (rowSums(costs.sub[,5:7]) / 3)

density.costs <- density(costs_vec, "SJ-ste")

density.costs

# Plot estimated distribution from 10,000 samples
set.seed(12345)
est.density.costs <- rkde(fhat = kde(costs_vec, h=0.07935), n = 10000)
hist(est.density.costs, breaks = 50, main = 'Estimated Value Distribution', xlab = "value")
```


2) Do one-year changes from 2006 to 2012 follow a normal distribution?

Since they are said to follow the same distribution as the 1990-2006 costs, let's look at the distribution of those (because the sample is larger). Based on the Q-Q plot below, those costs are normal, so we can assume the 2006-2012 costs are normal as well.
``` {r}
qqnorm( y = est.density.costs, main = "Normal Q-Q Plot",
       xlab = "Theoretical Quantiles", ylab = "Sample Quantiles",
       plot.it = TRUE, datax = FALSE)  
qqline(y = est.density.costs) # highlight and run with qqnorm 
```





3) Simulate possible future values of 2020 drilling costs under the assumption of normality as well as under the kernel density estimate you created (2006-2012) as well as the information from 2015 - 2020. Make a recommendation for which one you feel the company should use.

Under assumption of normality - just use the mean and standard deviation from the density estimate from 1990-2006

First calculating the average cost in 2006.

```{r}
P2006 <- rowMeans(costs[47,2:4])
```

Now going through simulating change each year to get 10000 simulations of 2020 cost

```{r}
set.seed(12345)

mu = mean(costs_vec)
stdev = sd(costs_vec)

norm_cost_2020 <- rep(0,10000)


for(i in 1:10000) {
  Pt <- P2006
  
  for(j in 2007:2020) {
    r <- rnorm(n=1,mean=mu, sd=stdev)
    Pt <- Pt*(1+r)
  }
  
  norm_cost_2020[i] <- Pt
}
```


Plotting distribution of possible 2020 costs with normality assumption
```{r}
hist(norm_cost_2020, breaks = 50, main = 'Estimated 2020 Cost Distribution - Normal', xlab = "Dollars")

mean(norm_cost_2020)
sd(norm_cost_2020)
```

Using info from kde, as well as 2012-2020.
- From 2006-2012, use sample from the kde estimate

- From 2012-2015, use a triangle distribution with mean -9.17%, min -22%, max -7%

- From 2015-2020, use a traingle distribution with mean 5%, maximum 6%, minimum 2%.

```{r}
set.seed(12345)
  
nonnorm_cost_2020 <- rep(0,10000)

for(i in 1:10000) {
  Pt <- P2006
  
  for(j in 2007:2012) {
    r <- rkde(fhat=kde(costs_vec, h=0.07935), n=1)
    Pt <- Pt*(1+r)
  }
  
  for(j in 2013-2015) {
    r <- rtriangle(n=1, a=-.22, b=-.07, c=-.0917)
    Pt <- Pt*(1+r)
  }
  
  for(j in 2016-2020) {
    r <- rtriangle(n=1, a=.02, b=.06, c=.05)
    Pt <- Pt*(1+r)
  }
  
  nonnorm_cost_2020[i] <- Pt
}

```


Plotting distribution of possible 2020 costs without normality assumption

```{r}
hist(nonnorm_cost_2020, breaks = 50, main = 'Estimated 2020 Cost Distribution - NonNormal', xlab = "Dollars")

mean(nonnorm_cost_2020)
sd(nonnorm_cost_2020)
```

We feel that the company should use the projected cost distribution that incorporates historical distribution from 1990-2006 as well as information known about 2012-2020 price changes. This will likely provide a more accurate estimate of 2020 prices, since the normal distribution projections do not take into account any information about the decrease in prices from 2012-2015