---
title: 'Simulation & Risk: homework 2'
author: 'Grant King'
date: "2/11/2020"
output: html_document
---
```{r packages}
library(car)
library(caret)
library(tidyr)
library(dplyr)
library(stats)
library(readxl)
library(ggplot2)
library(clipr)
library(ks)
library(EnvStats)
library(triangle)
library(knitr)
library(MASS)
library(rgl)
library(graphics)
```
This will run all of the hw1 code to get the data/objects. 
Paste the path to 'hw1_script.R' below..  
```{r setup, echo=FALSE}
opts_chunk$set(echo = FALSE, cache=FALSE)
read_chunk('C:/Users/grant/Documents/MSA/Spring/sim_risk/hw_2/hw1_script.R')
```

click green arrow to run this chunk
```{r hw1_results}
```

        
        --- 
 
        Step 1: Simulating the cost of a single dry well.

        ---    


Cost components:
  Drilling costs
  Seismic and Lease Costs
  Professional Overhead costs for year 0


```{r}
set.seed(1)
drill.cost.dry <- nonnorm_cost_2020 * 1000
lease.cost.dry <- rnorm(n=10000, mean = 600, sd = 50) * 960
seismic.cost.dry <- rnorm(n=10000, mean = 3, sd = .35) * 43000
prof.cost.dry <- rtriangle(n=10000, a = 172000, b = 279500, c = 215000)

dry.well.cost <- drill.cost.dry + lease.cost.dry + seismic.cost.dry + prof.cost.dry

hist(dry.well.cost)
```


        --- 
 
        step 2: Simulating the net present value of a single wet well

        ---  
        
Net present value components:
  Initial Costs:
    Drilling cost
    Seismic and Lease costs
    Completion costs
    Professional overhead costs for year 0
    
  Final net revenue for each year:
    Revenues - Costs
    Revenues:
      ((Oil price * Annual production) * NRI) * (1-severance tax)
      Oil price = triangle distribution for each year based on projections
      Annual production = calculated based on initial production and decline
      NRI = simulated, constant across well lifetime
      Severance tax= 4.6%
    Costs:
      Professional overhead each year - constant, simulated
      Operating costs - changes year to year
    

Initial cost
```{r}
drill.cost <- nonnorm_cost_2020 * 1000
lease.cost <- rnorm(n=10000, mean = 600, sd = 50) * 960
seismic.cost <- rnorm(n=10000, mean = 3, sd = .35) * 43000
completion.cost <- rnorm(n=10000, mean = 390000, sd = 50000)
prof.cost <- rtriangle(n=10000, a = 172000, b = 279500, c = 215000)

initial.cost <- drill.cost + lease.cost + seismic.cost + completion.cost + prof.cost
```

Yearly production

```{r}
volume_cumulative_30yrs <- rep(0,10000)
R <- matrix(data=cbind(1,0.64, 0.64, 1), nrow=2)
U <- t(chol(R))

standardize <- function(x){
  x.std = (x - mean(x))/sd(x)
  return(x.std)
}

destandardize <- function(x.std, x){
  x.old = (x.std * sd(x)) + mean(x)
  return(x.old)
}

set.seed(1)

IP.r <- rlnorm(n=10000, mean= 6, sd = 0.28)
decline.r <- runif(n=10000,min=.15, max=.32)
Both.r <- cbind(standardize(IP.r), standardize(decline.r))
IPD.r <- U %*% t(Both.r)
IPD.r <- t(IPD.r)
  
final.IP.r <- destandardize(IPD.r[,1], IP.r)
final.decline.r <- destandardize(IPD.r[,2], decline.r)

#changing decline to be within its regular bounds of .15 to .32, even though this will slightly mess with correlation
final.decline.r[final.decline.r<.15] <- .15
final.decline.r[final.decline.r>.32] <- .32


rate0 <- final.IP.r

#simulating declining rates for 30 years
for(year in 1:15){
  #set rate at end of year equal to rate at beginning minus decline
  #if the decline is more than 100% (due to correlation stuff) set rate=0
  rate1 <- rate0*(1 - final.decline.r)
  #calculate volume of oil in year
  eval(parse(text=paste('volume_',year,' <- 365 * (rate0 + rate1) / 2',sep="")))
  #set rate at beginning of next year to rate at end of previous year
  rate0 <- rate1
}


hist(volume_1, breaks=50)
hist(volume_5, breaks=50)
hist(volume_10, breaks=50)
hist(volume_15, breaks=50)

```

Yearly oil price

```{r}
for (year in 1:15){
  eval(parse(text=paste('price_',year,' <- rtriangle(n=10000, a=projections[[',year+1,',3]], b=projections[[',year+1,',2]], c=projections[[',year+1,',4]])',sep="")))
}
```

NRI, calculation of yearly revenue
```{r}
NRI <- rnorm(n=10000, mean=.75, sd=.02)
for (year in 1:15){
  eval(parse(text=paste('rev_',year,' <- (volume_',year,' * price_',year,') * NRI * (1-.046)', sep="")))
}
```


calculation of yearly costs

```{r}
for (year in 1:15){
  eval(parse(text=paste('op.cost_',year,' <- rnorm(n=10000, mean=2.25, sd=0.3)', sep="")))
  eval(parse(text=paste('op.cost.total_',year,' <- op.cost_',year,' * volume_',year, sep="")))
  eval(parse(text=paste('cost.total_',year,' <- op.cost.total_',year,' + prof.cost', sep="")))
}
```

calculation of yearly FNR

```{r}
for (year in 1:15){
  eval(parse(text=paste('fnr_',year,' <- rev_',year,' - cost.total_',year, sep="")))
}
```

calculation of net present value

```{r}

NPV = -initial.cost +
  fnr_1/(1.1) +
  fnr_2/(1.1^2) +
  fnr_3/(1.1^3) +
  fnr_4/(1.1^4) +
  fnr_5/(1.1^5) +
  fnr_6/(1.1^6) +
  fnr_7/(1.1^7) +
  fnr_8/(1.1^8) +
  fnr_9/(1.1^9) +
  fnr_10/(1.1^10) +
  fnr_11/(1.1^11) +
  fnr_12/(1.1^12) +
  fnr_13/(1.1^13) +
  fnr_14/(1.1^14) +
  fnr_15/(1.1^15)

hist(NPV, breaks=50)
mean(NPV)
```